---
title: "Boston House Prices Analysis"
author: "Kar Ng"
date: "2021"
output: 
  github_document:
    toc: yes
    toc_depth: 3
always_allow_html: yes
    
---


## 1 R PACKAGES

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(skimr)
library(caret)
library(MASS)

```

## 2 INTRODUCTION

This project uses a public dataset named "Boston" from the R package - MASS. This dataset studies the effects of a range of variables on median house prices in Boston near late 70s, United States. 

I will statistically analyse the dataset with machine learning algorithms to find out the effects of each variables on the median house prices and build a model that correctly predict the median house price.


## 3 DATA IMPORT

```{r}

Boston

```


## 4 EXPLORATORY DATA ANALYSIS (EDA)





## 


## REFERENCE 

Harrison, D. and Rubinfeld, D.L. (1978) Hedonic prices and the demand for clean air. J. Environ. Economics and Management 5, 81–102.

Belsley D.A., Kuh, E. and Welsch, R.E. (1980) Regression Diagnostics. Identifying Influential Data and Sources of Collinearity. New York: Wiley.



